Stack Safety as a Security Property

### 2020-07-08 main.tex
CP - intro

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion
is usually defined negatively, by enumerating the bad things that a given
enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one---quite recent---attempt to frame a
positive definition:
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
Although this represents an important step in the right direction, we believe
there is still room to craft a more abstract, more generic characterization
of stack safety that is independent of capabilities or any other specific
enforcement
technique.  In other words, we seek formal criteria that can be applied
against \emph{any} stack-based system + \rb{in words or symbols? :-)} enforcement mechanism to judge
whether or not the combination is actually stack-safe.

\iflater
\apt{another stab.}
\leo{I took a stab. I think it's still to dense, but captures the
  feeling of today's discussion I think}
\apt{still needs tuning}\bcp{Yes, I think we need to go into a
  little more detail on what their definition looks like in order to
  justify why an alternative definition might be valuable.  What
  questions does their presentation not answer? What uses does it not
  support? E.g., can it be enforced efficiently? (I guess so, by
  capabilities, though I think their capabilities are more powerful
  than what Cheri actually provides, while our policies are exactly
  what Nick needs.)  Why would it be good to look at stack safety
  through the lens of security?  (Because it is fundamentally a
  security mechanism?)}  \rb{BCP idea: contrast intrinsic
  vs. extrinsic property definition.  Remember to talk about spatial
  vs temporal safety}
\fi

The key contribution of this paper is a novel\bcp{generic?} presentation of
local state encapsulation and well-bracketed control flow using the
technical framework of language-based security.

The intuition behind \citet{Skorstengaard+19}'s\iflater\bcp{I forget how to
  remove the year and just have the author names here...}\fi use of the term {\em
  local state encapsulation} is that callers should be protected from
callees: a callee should be not be able to access or overwrite private data
in the caller's stack frame. We formalize this intuition as a security
property with both confidentiality and integrity components.
%
We reflect this intuition in a formal
{\em stack confidentiality} property, which promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g., a caller's private data or uninitialized memory).  Conversely, {\em stack
  integrity} promises that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.

@@

The intuition behind \citet{Skorstengaard+19}'s {\em well-bracketed control
  flow} is
the desire to protect the call graph itself: a callee should pass control
back to the caller only by returning in the standard way (transferring
control to the instruction following the call); it should
not be
allowed to jump back into the caller's code at an arbitrary point, either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety
between functions: each instruction is uniquely owned by some
function, and all jumps between functions are either calls that enter
at a marked entry point or returns that re-enter just after their
associated call in the standard nested call structure.  \apt{clarify
  what we are contributing here.}

To phrase these two aspects of stack safety in terms of security properties
requires an explicit attacker model. \bcp{This paragraph goes too fast: We
  need to start by asking, informally, who is being protected and who is the
attacker. Then we observe that real-world stack protection mechanisms tend
to imagine attack scenarios where the callee gets confused by something it
receives from the outside; to avoid getting into the details of how this
happens, we simply assume that the callee can execute arbitrary code.}
For simplicity, we choose one that is
fairly strong: when stating the protections that a function can rely on when
it makes a call, we assume that the callee can execute arbitrary
instructions until it returns.

>>>>

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion
is usually defined negatively, by enumerating the bad things that a given
enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one---quite recent---attempt to frame a
positive definition:
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
Although this represents an important step in the right direction, we believe
there is still room to craft a more abstract, more generic characterization
of stack safety that is independent of capabilities or any other specific
enforcement
technique.  In other words, we seek formal criteria that can be applied
against \emph{any} stack-based system + enforcement mechanism to judge
whether or not the combination is actually stack-safe.

\iflater
\leo{Should this be its own paragraph?}
\fi
The key contribution of this paper is a novel, generic presentation of
local state encapsulation and well-bracketed control flow using the
technical framework of language-based security.

The intuition behind \citeauthor{Skorstengaard+19}'s use of the term {\em
  local state encapsulation} is that callers should be protected from
callees: a callee should be not be able to access or overwrite private data
in the caller's stack frame. We formalize this intuition as a security
property with both confidentiality and integrity components.
%
We reflect this intuition in a formal
{\em stack confidentiality} property, which promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g., a caller's private data or uninitialized memory).  Conversely, {\em stack
  integrity} promises that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.

@@

The intuition behind \citeauthor{Skorstengaard+19}'s {\em well-bracketed control
  flow} is
the desire to protect the call graph itself: a callee should pass control
back to the caller only by returning in the standard way (transferring
control to the instruction following the call); it should
not be
allowed to jump back into the caller's code at an arbitrary point, either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety
between functions: each instruction is uniquely owned by some
function, and all jumps between functions are either calls that enter
at a marked entry point or returns that re-enter just after their
associated call in the standard nested call structure.  \apt{clarify
  what we are contributing here.}

To phrase these two aspects of stack safety in terms of security
properties \leo{in terms of security properties or ``using security
  language''?}  requires an explicit attacker model: an clear
understanding of who is being protected and from what. Most existing
stack protection mechanisms deal with scenarios where a callee gets
influenced by external inputs and ends up executing potentially
malicious code. To avoid the details of exactly how this happens and
keep the presentation clear, we pick an even stronger model: when
stating the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.
